Implications of uniformly distributed, empirically informed priors for phylogeographical model selection: A reply to Hickerson et al.

### 2014-07-30 manuscripts/reply/reviews/response.tex

%&<latex>
\documentclass[12pt]{article}

\usepackage{parskip}
\input{../../utils/preamble.tex}
\usepackage[numbers,sort&compress]{natbib}
\input{../../utils/macros.tex}

\newcounter{commentCounter}
\newcommand{\revcomment}[1]{{\addtocounter{commentCounter}{1}}
    \medskip \hrule \noindent
\textbf{\arabic{section}.\arabic{commentCounter}}: {\sl #1}\par\xspace}
\newcommand{\response}[1]{{\addtolength{\leftskip}{0.25in} #1\par}\xspace}

\let\quoteOld\quote
\let\endquoteOld\endquote
% \renewenvironment{quote}{\sffamily\slshape\small\quoteOld}{\endquoteOld}
\renewenvironment{quote}{\sffamily\small\quoteOld}{\endquoteOld}

% \makeatletter
% \renewcommand{\section}{\@startsection{section}{1}{0mm}%
%     {-8pt}%
%     {4pt}%
% {\sffamily\large\itshape}}
% \makeatother

\begin{document}
% \raggedright
\setlength{\parindent}{0in}

I greatly appreciate the helpful comments of the editor and reviewers. In this
document, I provide responses to these comments, including descriptions of the
changes I have made.

\section{Editor's Comments}\label{ed}
\setcounter{commentCounter}{0}
\revcomment{
    The most negative review suggests that the authors (and Hickerson et al)
    might have reached a point of diminishing returns, where further exchange
    in the pages of Evolution might not be informative for our readership. My
    own view is somewhat in line with this reviewer. I do think there are some
    new ideas in the present manuscript, and I also want to honor the previous
    AE who solicited an additional response from Oaks and colleagues to the
    last Hickerson et al. paper. But I would also say that after this paper,
    enough is enough, in my opinion \ldots
}
\response{
    We agree. We are very appreciative of this opportunity to clarify some
    aspects of Hickerson et al.'s response, and feel our reply will make a nice
    end to an interesting discussion about Bayesian model selection.
}

\revcomment{
     I also echo the suggestion from reviewer 2 to shorten this paper as much
     as is practical, especially by moving things to the supplement. One can
     safely assume that readers still following the thread have read the
     preceding papers in the series, I think.
}
\response{
    We are greatful for this insight by the editor and reviewers that our
    paper was too long and lacked focus.
    We have greatly reduced the length of the main text and moved some items
    (including 2 figures) to the supplement.
}

\revcomment{
    One small point: there is a typo on line 19: Oaks (2014) should be (Oaks
    2014).
}
\response{
    Corrected. Thank you.
}

\section{Reviewer 1's Comments}\label{rev1}
\setcounter{commentCounter}{0}
\revcomment{
    At this point, I think at least for the readers of Evolution, we are
    reaching the point of diminishing returns and perhaps the discussion has
    burned out. The arguments of how broad priors or mixes of specific and
    diffuse priors yield poor posteriors is indeed important but perhaps
    another venue would be appropriate for this discussion---also with a
    greater emphasis on solutions, if they exist for non-model organisms, to
    common problems affecting most molecular ecologists (particularly regarding
    substitution rate uncertainty).

}
\response{
    We agree that Bayesian statisticians are aware of the issues raised in our
    paper. However, I think among biologists there is common misunderstanding
    about some of the subtle but fundamental differences between Bayesian
    approaches to parameter estimation and model selection. Mainly, the
    implicit and strong statements that priors on the models' parameters can
    make about the probability of the models. Thus, we feel the readership of
    \emph{Evolution} will benefit from issues discussed in the paper.

    We have added emphasis of solutions to the problem in several places in the
    manuscript (abstract, introduction, and conclusions). The solution was the
    focus of a different paper \citep{Oaks2014dpp} that introduces an improved
    method for estimating shared divergence histories. We now draw attention to
    this positive outcome from this work.
}

\section{Reviewer 2's Comments}\label{rev2}
\setcounter{commentCounter}{0}

\revcomment{
    The paper is rather long for simply arguing against a model averaging
    approach. I recommend to move the coin example to the supp info, as this is
    a simple text book example that does require to be published in a journal
    such as evolution, but is still a nice illustration so that it should be
    published as supp info to the paper. Also, I think the paper sometimes
    reads like a direct accusation of the findings in Hickerson et al, to agree
    to which they might be more emotional than required for the presented
    arguments. Paragraphs such as L96-100, L104-108, L222-237 or L287-289
    (among others later in the text and, especially, in the supp info) for
    instance, could simply be omitted or shortened, which would make reading
    the paper much more enjoyable, as it would focus on the valid arguments,
    instead of highlighting subtleties only personally important to people
    included in the debate.

}
\response{
    We thank Reviewer 2 for this insight. We agree that the paper was too long
    and lacked focus. Reviewer 2 also points out several sections that might be
    interpreted as emotionally charged or polemical, which we appreciate. We
    did not intend for our response to read as an ``attack'' of the work of
    Hickerson et al., but Reviewer 2 helped us to realize that wording in a
    response paper needs to be carefully chosen to avoid this interpretation.

    We have moved the coin-flipping example to the supplement.

    In regard to the sections specifically highlighted by Reviewer 2:
    \begin{description}
        \item[L96-100 and L104-108]
            We have reduced this section to only three sentences that summarize
            only the information necessary to introduce our subsequent
            discussion.
        \item[L222-237]
            We see how this paragraph can be interpreted as polemical. However,
            this brief summary of the methods and results of Hickerson et al.'s
            model-averaging approach is needed to set up the motivation for our
            subsequent analyses.
        \item[L287-289]
            We agree this section was superfluous. We have removed it.
    \end{description}

    We have also removed and edited many other sections of the paper to try to
    eliminate sections that might seem overly negative. Specifically, we
    removed all quotations of sections from Hickerson et al. While we only
    intended to accurately represent their positions, we felt this is too
    easily interpreted as contentious.
}

\revcomment{
    While criticizing the approach proposed by Hickerson et al. and admitting
    the power problem of msBayes, the paper falls short of providing much of an
    alternative. However, I do think that there could be other ways of solving
    the problem of too many parameters for few summary statistics (section 6).
    This is especially true since ABC is known to be particularly sensitive to
    the choice of summary statistics for model choice (see Roberts et al. PNAS
    2011). I think the paper would be much stronger, if it could also discuss
    (but not necessarily test or implement) potential solutions to the apparent
    problem.
}
\response{
    This is a great point, and we have added emphasis to solutions to the
    problem, as recommended by Reviewer 2. As mentioned above, solutions to
    this problem were the focus of a separate paper, and were too extensive to
    discuss in detail in the response paper. However, we highlight the main
    points of this paper to make the response more positive, and to to refer
    readers to an improved methodology.
}

\revcomment{
    One alternative suggestion could be to formulate the problem as a
    hierarchical model, in which one seeks only to infer mean and variance in
    divergence times. Such a model would be much easier to fit, as the summary
    statistics ordering problem discussed (and solved) in the original msBayes
    paper could be circumvented easily, as the model would not seek to infer
    taxon specific divergence times. Also, it could then be compared to a model
    of zero variance using either Bayes factors, or by assessing the posterior
    probability of the variance to be zero using a simulation / bootstrap
    approach. If a model of co-divergence is preferred, the inferred posterior
    on the mean divergence would directly serve as an estimate of the joint
    divergence time. If independent divergences are inferred, then a simpler
    model of a single taxon pair could be used to infer the divergence times of
    each taxon pair individually, as they were found to be independent. For
    such a model, the same ABC simulations could be reused for all pairs,
    making this step computationally very cheap. But I'm sure there would also
    be other alternatives.
}
\response{
    These are great points raised by Reviewer 2. Many of these are
    discussed in detail in a separate paper \citep{Oaks2014dpp}.
    We highlight the improved methodology presented there to add a positive
    note to the response.
}

\revcomment{
    A minor issue on L53: while it is true that all numerical methods suffer
    from large prior ranges, it is also true that effect is pronounced in ABC,
    as the sampling occurs from the prior, and not the posteriors, as is done
    in techniques such as MCMC. Hence, it is not a problem that is completely
    extrinsic to the ABC machinery.
}
\response{
    We thank Reviewer 2 for pointing out this section, which was admitedly very
    confusing. We have re-written this section (and almost all of the
    introduction) to try to make this distinction clearer. Our goal is to
    distinguish two different ways that broad uniform priors could cause the
    poor behavior of msBayes that was shown in the origin Oaks et al. paper.
    \begin{enumerate}
        \item As pointed out by reviewer 2, these priors could cause numerical
            issues, where the simple ABC rejection machinery is unable to
            sufficiently sample the space of the model in manageable
            computational time.
        \item The priors can cause rich models to have very small (but correct)
            marginal likelihoods.
    \end{enumerate}

    These issues are not mutually exclusive, but in the former the true
    posterior would support heterogeneity in divergence times in the face of
    random divergences (i.e., the problem only arises from our inability to
    estimate that posterior).
    Whereas in the later, the true posterior (i.e., an analytical solution)
    would correctly tell us that the most probable model is simultaneous
    divergence despite random variation in divergence times. I.e., this is
    actually the correct answer under the prior assumptions from the
    perspective of Bayesian model choice. Unfortunately, this correct answer
    does not provide the biogeographical insights we are using the method to
    gain.

    This is an important distinction, because if the prior conditions are such
    that we are in situation 2, any arguments about the estimation procedure
    are moot, because even the true posterior is biogeographcially misleading.
}

\bibliographystyle{../../bib/bmc-mathphys.bst}
\bibliography{../../bib/references}

\end{document}

### 2014-07-30 manuscripts/reply/reply-abstract.tex
More updating of reply ms.
Finishing draft of reply ms overhaul.

% Biogeographers frequently seek to explain population and species
% differentiation on geographical phenomena.
Establishing that a set of population splitting events occurred at the same
time can be a potentially persuasive argument that the populations were
affected by the same geographic event.
% Unfortunately, estimating divergence times precisely when one lacks
% precise information about the rate of molecular evolution is a notoriously
% difficult problem in evolutionary biology.
% \citet{Huang2011} introduced an approximate-Bayesian approach, \msb, for
% estimating the probabilities of models in which multiple sets of taxa diverge
% at the same time.
% This method  does not require a common set of character data that links all
% of the taxa under investigation, but does require the relative mutation
% rates among the taxa and loci to be assumed \emph{a priori}.
Recently, \citet{Oaks2012} used an approximate-Bayesian model choice method
(\msb) to estimate the support for such a pattern of divergences across 22
pairs of taxa from the Philippines.
They also studied the behavior of \msb using computer simulations.
\citet{Oaks2012} found the method was very sensitive to prior assumptions and
had low power to detect variation in divergence times.
These results were not surprising in light of a rich statistical literature
showing that the marginal likelihood of a model is sensitive to the
use of vague priors.
Because this sensitivity to prior assumptions affects the crucial insights that
a researcher who employs \msb seeks to gain, \citet{Oaks2012} recommended that
users of the approach should carefully assess the robustness of their
conclusions to different priors.
In response, \citet{Hickerson2013} suggested the lack of robustness in \msb
analyses was due to the method's inefficient rejection algorithm and broad
priors on divergence times leading to inadequate numbers of simulation
replicates.
They proposed a model-averaging approach that uses narrow, empirically
informed uniform priors.
Here we demonstrate that the approach of \citet{Hickerson2013} will often
incorrectly support shared divergence events and is dangerous in the sense that
the empirically-derived priors often exclude from consideration the true values
of the models' parameters.
On a more fundamental level, we question the value of adopting an empirical
Bayesian stance for this model-choice problem.
% , because it can mislead model
% posterior probabilities, which are inherently measures of belief in the models
% after prior knowledge is updated by the data.
The robust Bayesian approach of conducting analyses under a variety of priors
can reveal prior sensitivity and communicate which assumptions underlie model
inference.
Furthermore, simulations provide insight into the temporal resolution of the
method, which helps guide interpretation of results.
Fortunately, as predicted by fundamental principles of Bayesian model choice,
more appropriate priors on the models' parameters can accommodate prior
uncertainty while increasing the method's robustness and power to detect
temporal variation in divergences
\citep{Oaks2014dpp}.

>>>>

Establishing that a set of population splitting events occurred at the same
time can be a potentially persuasive argument that the populations were
affected by the same geographic event.
Recently, \citet{Oaks2012} used an approximate-Bayesian model choice method
(\msb) to estimate the support for such a pattern of divergences across 22
pairs of taxa from the Philippines.
They also studied the behavior of \msb using computer simulations.
\citet{Oaks2012} found the method was very sensitive to prior assumptions and
had low power to detect variation in divergence times.
These results were not surprising in light of a rich statistical literature
showing that the marginal likelihood of a model is sensitive to the
use of vague priors.
Because this sensitivity to prior assumptions affects the crucial insights that
a researcher who employs \msb seeks to gain, \citet{Oaks2012} recommended that
users of the approach should carefully assess the robustness of their
conclusions to different priors.
In response, \citet{Hickerson2013} suggested the lack of robustness in \msb
analyses was due to the method's inefficient rejection algorithm and broad
priors on divergence times leading to inadequate numbers of simulation
replicates.
They proposed a model-averaging approach that uses narrow, empirically
informed uniform priors.
Here we demonstrate that the approach of \citet{Hickerson2013} will often
erroneously support shared divergence events and is dangerous in the sense that
the empirically-derived uniform priors often exclude from consideration the
true values of the models' parameters.
On a more fundamental level, we question the value of adopting an empirical
Bayesian stance for this model-choice problem.
The robust Bayesian approach of conducting analyses under a variety of priors
can reveal prior sensitivity and communicate which assumptions underlie model
inference.
% Furthermore, simulations provide insight into the temporal resolution of the
% method, which helps guide interpretation of results.
As predicted by fundamental principles of Bayesian model choice, more flexible
distributions can accommodate prior uncertainty about the models' parameters
while increasing the method's robustness and power to detect temporal variation
in divergences \citep{Oaks2014dpp}.
